---
title: "Research Note: Why and When do Leaders of Developing Countries Negotiate South-North Preferential Trade Agreements?"
author: "Lucie Lu"
date: "October 20, 2018"
output: pdf_document
fontsize: 12pt
geometry: margin = 1.2in
header-includes:
- \usepackage{placeins}
- \usepackage{setspace}
- \usepackage{chngcntr}
- \usepackage{pdflscape}
- \onehalfspacing
- \counterwithin{figure}{section}
- \counterwithin{table}{section}
- \usepackage{float}
- \floatplacement{figure}{H}

bibliography: reference_fyp.bib

---

```{r global_options, include=FALSE, cache=FALSE}
## To make the pdf file do
## render("exploration4.Rmd",output_format=pdf_document())

require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",    # slightly smaller font for code
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132,
  size='footnotesize',
  out.width='.9\\textwidth',
  message=FALSE,
  comment=NA)

options(knitr.graphics.auto_pdf = TRUE)

```


```{r initialize,echo=FALSE, message=FALSE, warning=FALSE}
##First, just setup the R environment for today:
if(!file.exists('figs')) dir.create('figs')

options(SweaveHooks=list(fig=function(){
			   par(mar=c(3.5, 3, 1.1, 0),
			       pty="s",
			       mgp=c(1.5,0.5,0),
			       oma=c(0,0,0,0))},
			 echo=function(){options(continue=" ") ##Don't show "+" prompts,
			 options(prompt=" ")
			 }),
	digits=4,
	scipen=8,
	width=132
	)
options(error=function(){options(prompt="> ",continue="+ ");NULL})


#To prepare the environment
library(readstata13)
library(dplyr)
library(ggplot2)
#install.packages("ggrepel")
library(ggrepel)
library(stargazer)

#install.packages("devtools")
library(devtools)

library(foreign)
#install.packages("gplots")
library(gplots)

library(lmtest)
library(sandwich)

suppressMessages(library("tidyverse"))

library(MASS)
library(robustbase)

#install.packages("rmngb")
library(rmngb)
library(here)

#devtools::install_github("ropenscilabs/gramr")
library("gramr")
#write_good_ip()

library(arm)
#install.packages("effects")
library(effects)
library(ROCR)

library(multiwayvcov)

#devtools::install_github("rstudio/rmarkdown")
#install.packages("stringi",type="win.binary")

#install.packages("kableExtra")
#install.packages("magrittr")
library(kableExtra)
library(magrittr)
knitr::opts_chunk$set(echo = TRUE)

```


	
$\qquad$ In 1994, South Africa took an opportunity to negotiate a preferential trade agreement with its most important economic partner in history, the European Union (EU). The same year also featured the victory of Mandela and the African National Congress (ANC) from the first universal suffrage elections. However, this new government was measured as one of the most extremely unstable and insecure ones. Interestingly, the ANC demonstrated a dramatic shift in relation to the policy stance on privatization. It had consistently opposed the competitor's economic initiative of privatization and deregulation, and advocated a state-led developmental plan and even threaten re-nationalization of privatized assets [@habib_economic_2000]. Such opposition to privatization evaporated when Mandela and the ANC won the election. Mandela's government wanted to conduct substantial privatization and other neoliberal economic reforms to restore the collapsing economies, but the reason of the government's dramatic shift in the economic strategy remained unknown. This new government's initiatives were strongly opposed by the other influential domestic political players. Mandela immediately approached to the EU to start the process of Preferential Trade Agreement negotiation. After five years of negotiations, the preferential trade agreement with the EU helped South Africa to undergo its transitioning economies smoothly.  

$\qquad$ South Africa's experience is a successful case for developing countries to pursue its reform-oriented trade strategy in the negotiation of PTAs with the major trade powers. Economic reforms usually provoke controversy and organized opposition domestically, so often the ability of leaders to implement such liberal economic reforms is limited [@bacciniInternationalInstitutionsDomestic2014]. Preferential trade agreements provide member governments with a mandate to make policy changes, as such they entail high political costs. Recognizing the preferential trade agreement with provisions of economic reforms is a binding commitment that imposes policy reforms and entails political costs, some leaders in the developing countries chose to negotiate it with the major liberal trade powers, while some did not. My research question is: Why and when will leaders in the developing countries negotiate South-North preferential trade agreements with provisions of economic reforms?  


$\qquad$ This paper is interested in explaining what motivates a leader from the developing countries to negotiate this particular deep and demanding international treaty. Why is this an important question? One reason is the popularity of this type of preferential trade agreement. The bilateral preferential trade agreement is a rapidly growing international institution that governs economic markets of the member states. Leaders are more interested in building bilateral trade agreements because it is easier to realize both parties' initiatives to achieve agreements when the participants are limited during the negotiations. South-North and South-South agreements are also growing: from the 1990s onward, they comprise two-thirds of the total PTAs in force [@worldtradeorganizationWTOPreferentialTrade2011]. In addition, PTA tends to have deeper integration, a trend especially prominent in the South-North PTA. They tend to go beyond tariff reductions but include provisions on regulatory policy areas such as foreign investment, intellectual property rights and human rights issues. This paper focuses on the PTA with the provisions on the competition policy as a stringent requirement imposed on contracting countries to ensure the structural reforms. The provision includes but is not limited to the privatization of the state-owned firms and regulation on monopolies and cartels. In other words, these PTAs are designed to change the domestic economic structures and reset the rules of the game for the contracting states who have not met the standards yet. This specific type of South-North PTAs this paper focuses on is one of the most costly and deepest ones because it requires the states to change their policies, conduct structural economic reforms and open its domestic markets for competition.  


$\qquad$ This paper also intends to focus only on the leaders from the South and their rationales of initiating a negotiation with the North. I assume the leaders from the developing countries are fully aware of the costs of complying with these international agreements, yet they still choose to do so. The power imbalances between the contracting parties are particularly stark in the South-North PTAs negotiations, reflecting the different potential costs for them. Developing countries make a larger concession to play the rules of the game mainly set by the major liberal players. In other words, when they approach these major powers to negotiate a PTA, they more or less anticipate they have chosen the deepest PTAs by design. Existing research tends to downplay this perspective, and neglect how international institutions enter leaders' political calculus differently. Yet, this is an important issue because leaders' choices reflect their motives of negotiating a PTA at a specific time. It can help us getting closer to contribute to a broader research debate: when leaders commit to the international agreement, is it because the agreement change their behaviours or they were planning to do so anyway before the commitment? The South-North PTA negotiation that encompasses the power imbalances between the two negotiating parties illuminates why leaders commit to international agreement in general. Leaders from the South make a larger concession to play the rules of the game set by the North, so the leaders have chosen the most costly game by design. By studying the choices leaders make, we get a step closer to understand their motives. 



#Context and Rationale

$\qquad$ Preferential trade agreements (hereafter PTAs) are binding international treaties that help states to foster trade and economic integration among member-states. Current literature on the preferential trade agreement mainly focuses on the effects of this institution. Questions centre around whether the PTAs increase trade volumes among member-states and their impacts on the overall trade flows in the multilateral trade system [@bagwellWillPreferentialAgreements1998]. Put simply, scholars examine to what extent the PTAs have achieved their desired economic purposes: lowering tariffs, facilitating trade and increasing the welfare of the states in the long term. In addition to studying the consequences of the PTAs, emerging political economy literature has offered explanations for why states form the PTAs. Earlier years, the focus is the macroeconomic and macropolitical factors that explain the proliferation of the PTAs, including the lack of the progress in the multilateral trade regime and the spillover effect of regionalism [@baldwinDominoTheoryRegionalism2012; @rodrikRushFreeTrade1992; @mansfieldMultilateralDeterminantsRegionalism2003]. 


$\qquad$ One trend of the current literature on the PTA focuses on the effects, others on the cause, which is the theme that this paper speaks to. Recent studies zoom into the domestic factors within the states to explain this trend. While there is no shortage of explanations on why governments form PTAs, scholars also tend to treat the PTAs as a unitary concept [@durDesignInternationalTrade2014]. They overlook the design differences of the PTAs, which imply different degrees of policy changes embedded in these international agreements. Few studies reach beyond trade liberalization to examine the role of the PTA in policy reform. Perhaps the most important exception is Baccini and Urpelainen [-@bacciniInternationalInstitutionsDomestic2014], who argue PTA can help developing countries to implement economic reforms, leaning more on examining the effect of the PTA. Picking up on this approach, this article argues the leaders' commitments to such costly agreement reflect their political purposes. Leaders have political objectives to commit to the deep PTAs with the expectation to conduct economic reforms after they face political challenge. The political motivation of PTA negotiation is currently understudied; yet, to understand when and why a leader of developing countries commits to the preferential trade agreement with the provision of economic reform is important because economic reforms will have immense consequences for the citizens of that country and beyond. The key question, the cause of economic reforms, yet, remains open.


$\qquad$ To explain why leaders sign trade agreements, the existing theoretical framework only focuses on the following domestic actors: median voters, interest groups or private sectors. This article is in line with the argument proposed by Mansfield and Milner [-@mansfieldVotesVetoesPolitical2012]: "government sign PTAs for domestic political reason" (p. 23). They offer one potential explanation: the regime type. Democratic countries are more likely to join a PTA than nondemocratic countries because these governments can send a positive signal of their liberal trade policies commitments to the median voters [@mansfieldVotesVetoesPolitical2012; @mansfieldWhyDemocraciesCooperate2002]. New leaders in democratizing countries tend to negotiate the PTA to lock in economic reforms to avoid autocratic reversals [@bacciniInternationalInstitutionsDomestic2014]. To understand why a government negotiates the PTAs with other government, scholars also focus on the interest groups who benefit from the expanding market promised by the PTAs. The pro-free trade beneficiary groups lobby the incumbent governments to sign such a trade agreement to serve their political objectives [@grossmanPoliticsFreeTradeAgreements1995]. On the other hand, trade policy affects groups in society differently. Some private sectors benefit from the protectionist policy. Over time, they will expand too much, create rents and inefficiencies to distort the economy. Governments make international commitments in the free trade international agreements to tie their hands to avoid making an excessive concession to interest groups in the expectation of protection and subsidies in the future [@staigerGattRulesHelp1999]. These arguments capture the interaction between a government and its domestic audiences: median voters, interest groups or private sectors. Within these frameworks, leaders generally use the international trade agreements to signal the domestic audiences of their welfare-enhancing economic policy. I argue it is deficient to see their motivation as economically driven only. I find a key player is missing: the opposition. What is overlooked is that leaders negotiate PTAs after they experience political crisis. To advance the literature on understanding the cause of PTA formation, this paper focuses on the strategic interactions between the leader and the opposition.   


$\qquad$ My argument is, leader negotiates PTAs after political crisis to punish the opponents' past transgression. PTA negotiation is a deliberate choice that a leader makes to commit to the binding and stringent policy reforms to rearrange the domestic power structures.


#Theory

##Economic reform as a political strategy

$\qquad$ I argue PTA negotiation with the stringent provision of economic reform is a tactic a leader uses to reduce the strength of opposition after he survives a political crisis. To understand why this argument holds, I first lay out the basic set-up in a non-crisis circumstance: a leader offers rents to exchange for the opposition's support. This arrangement does not prevent a political crisis, which can happen at any time. The opposition group attempts to depose the leader, but this group fails. The leader survives the shock to security, so now he has two tactics to punish the defectors: further repressing or retracting loyalty rents. Economic reform is a powerful tactic to rearrange the power bases of the elites, so it serves as a punishment to the defectors.   

$\qquad$ Let us start with the basic setup. Leaders' most fundamental interest is survival. They inherently want powerful and stable leadership. Leaders have two instruments to accumulate power: repression and loyalty [@wintrobe_political_1998]. Repression means a leader restricts individuals' rights to criticize against the government or use severe sanctions to punish the organized opposition to the government, including imprisonment, execution, and torture. Loyalty means a leader can accumulate the political loyalty of elites by means of the distribution of rent. Political rent is the price a leader pays to buy off the loyalty of the opposition to exchange for political stability. Certain economic policies create rents for elites that serve as the bases of supports. Heavy subsidies to State-Owned Enterprises (SOEs) and import-substituting industrialists, and political project of the government are some examples of policies of loyalty. The continuing loyalty of elites is critical to leaders' survival.    


$\qquad$ Opponents compete for power. Some have independent power to consider certain types of legislation; some can form political parties to compete in office formally, and some may share covert power bases with interest groups in society. Because a leader's power is not absolute, opposition may defect and overthrow the incumbent government from time to time. Their ultimate goal is to subvert the rule of the current leader. They can choose to form a coalition with the leader or pose a threatening demand to the current regime.   


$\qquad$ In non-crisis circumstances, leader and opposition already have mutual signalling and commitment problems. To reward the loyalty of the opposition, a leader could promise to implement certain policies that the opposition benefit. One example is the government's subsidies to the SOEs or some public financed projects that create rents for the opposition. The opposition, benefiting from such an arrangement, offers support for the regime in return. However, what guarantee does each party have that the other party will live up to their commitment? Leaders could potentially see the risks of increasing power of the opposition who benefit from collecting political rents, so they could stop appeasing the opposition but start torturing them. The opposition may also promise loyalty to the leader on the surface, but they could take part in an assassination plot as soon as the ruler shows his weaknesses. In sum, the so-called Dictator's Dilemma always exists: "the greater the dictator's power, the more reason he or she has to be afraid" [@wintrobe_political_1998, p. 22]. Stated differently, a leader never really knows how popular he is and when the promised support stops being delivered. No matter how stable a regime may look, an upheaval may occur any time [@kuranNowOutNever1991].   

$\qquad$ The theoretical story begins with the post-crisis circumstances: a shock to security occurs and the leader survive the shocks. We can conceptualize shock to security as "the [threatening] demands on a regime by opposition groups" [@reganDemocracyThreatsPolitical2002, p. 122]. A shock to security occurs when opponents pose threatening demands. The opponents' abrupt demands can take various forms, such as strikes, protest demonstrations, riots, guerrillas and civil war. The ruler will perceive the abrupt demands as an open challenge from the opposition and associate them with defecting from the promised loyalty. In this study, a *shock to security* occurs when opponents pose threatening demands to challenge the current regime, which leads to political instability.   


$\qquad$ The story focuses on what leaders behave in the post-crisis situation. What is the leader's response after the opposition defect and pose threatening demands? The most intuitive response is the leader will punishes them. If the costs of reneging are high, the frequencies of defection are expected to drop. Scholars mainly focus on the violent forms of political repression as the leader's main response to deter opposition: political repression. I argue a different tactic a leader can use, a non-violent form of punishment: economic reform. This adds more nuances to the instrument of a leader uses in the loyalty side: he can retract the promised loyalty to rearrange the power bases.   


$\qquad$ One common route a leader takes is to torture or murder the opponents. I argue repression is only one of the tactics a leader can use to punish the defectors. A leader who survives a shock may not have the capacity to kill all the revolutionary leaders and the conspirators in the organized opposition groups as he wishes. Even if he has the capacity to do so, all the opposition today will become his potential threats in the future, so it is not feasible to kill all of them. Therefore, repression is one of the tactics that a leader uses; however, it may be not the wisest one.   


$\qquad$ I argue, another powerful tactic of punishment is economic reform. This strategy withdraws the promised payoffs and limits the rent-seeking opportunities for the opposition today and tomorrow. Economic reforms, i.e., the privatization of state-owned enterprises and the open and fair competition for the government-funded project can effectively cut off the power sources of the opposition. Prior to the crisis, all the opposition somehow benefit from the rent. After the crisis, it is time for the leader to rethink his pre-crisis rent distribution. Now, economic reforms come at the expense of the interest groups who previously privileged from the political rents. Economic reforms can cut off the power bases of the elites who previously privileged from the political rent. The structural changes of economies also make the opposition's future accumulation of rents and power harder. I Economic reform is a powerful tactic to punish the defectors.


##Why negotiating a PTA with major liberal powers?

$\qquad$ Suppose a leader perceives economic reform as a tool to punish the opposition and announces his decision to conduct an economic reform domestically without a PTA negotiation. Earlier on, we have discussed the commitment problem among the leader and opposition is acuter in autocracies and immature democracies. The domestic opposition are skeptical about the leader's credibility and capacity to implement such a policy. Especially after the shock to security, a leader needs time to restore the status quo ante yet strong enough to implement the reform unilaterally. Failure to implement economic reform will not only by no means serve as a punishment to the opposition and deter future dissent, but also diminishes the credibility of leaders' future policy changes.   


$\qquad$ PTAs with major liberal powers mitigates the commitment problem highlighted above. The negotiation of PTA sends a credible signal to the foreign investors that economic reforms will be more likely to be enforced. Recall, PTAs are binding international agreement: governments cannot constantly violate or arbitrarily pull out of the agreements without retaliation. PTAs by design have some coercive mechanisms of influence to make the costs of defection higher than the costs of conformity. In plain words, the expected values of the preferential trade agreements are conditional on the implementation of economic reforms. If leaders fail to implement the provisions, they will pay costs. Therefore, leaders in the developing countries approach to those in the developed countries to negotiate a stringent PTA implies that leaders have anticipated complying the provisions of economic reforms. PTA negotiation sends a credible signal of leaders' commitments to the reforms.   


$\qquad$ PTAs negotiation also sends a strong signal to the opposition to reveal the leader's capacity to conduct such reform after the negotiation. While a leader's announcement of economic reforms will probably induce backfiring by the opposition, the successful negotiation of PTA signals that the power of the current leader is strong. The leader has probably recovered from the previous crisis and has the capacity to conduct the reform. This can reduce the possibilities that the opposition overthrow the leader again after their previous failed attempt. Hence, PTA increases the chance of opposition's cooperation with the reform.
 


$\qquad$ Now let us turn to explain why PTAs can help leaders to implement economic reforms. The PTAs are expected to increase trade volumes and the flow of foreign direct investment, and hence increase tax revenues. As a result, the leader has more resources at their disposal, furthering their hold on power to recover from the shock to security. The promising trade benefits also expect to increase national income. Leaders can claim credit for the improvement in welfare, so the citizens' loyalty to the leadership will be higher. The opposition, who mainly lose from the liberalized economic reforms, can be pacified by the values of the PTAs. Now with the PTA, a leader can use the increased trade revenues as a side payment to compensate the loyal opponents. Alternatively, leaders can further hit the disloyal opponents by deeper reforms in certain economic sectors. With the PTA, a leader not only uses economic reforms to target the disloyal opponents, but also reestablishes a coalition with loyal opponents to consolidate his power.   


##Summary 

$\qquad$ The puzzle to explain is what motivates a leader in the developing country to negotiate one of the most costly and stringent trade agreements with the developed country. One of the most important features of these PTAs is the provision of economic reforms, whose implementation will ignite domestic opposition. My argument, in brief, is the following: a leader strategically negotiates a PTA when he is insecure. I argue after a leader is hit by a shock to security, he can choose economic reforms to punish the disloyal opposition responsible for the outbreak. Leaders are more likely to implement economic reform when the PTAs negotiation is in place. The PTAs with major economic powers send positive signals to mitigate commitment problem that a leader has, and offers material benefits to facilitate the reforms. This argument has a main empirical implication that can be used to test its validity in the next section. My prediction is: Negative shocks to a leader's security increase the probability of PTA negotiations.    


#Research Design


$\qquad$ My research design compares the likelihoods of leaders with shocks to security and those without shocks to negotiate the South-North PTAs. The hypothetical treatment in this study is *shock to security*. I defined two types of shocks to security. Before getting into a shock, here I use two indices to measure the security of a leader:   

1) a leader's security before he starts his tenure at time *0* depending on his relation to the past leader (or not); and,  

2) the security of the regime when the leader holds office at time *t*.     


$\qquad$ At time *0*, regime type matters. I use Archigo's data set on leaders to obtain information of state leaders' names, their tenures, the regime types when they hold power, manners of leader transfers [@gleditsch_archigos_2015]. The regime types are divided in three cateogires: democracy, autocracy and no authority. They are identified based on whether political leaders enter and exit in political office based on explicit rules. V-Dem's Electoral Democracy Index [-@coppedge_v-dem_2018] is also used to cross validate the regime type measurement. 

$\qquad$ An authoritarian leader is coded as secure at time 0 when he is politically affiliated and from the same ruling coalition with his previous leader. Otherwise, he is insecure at time 0. To measure leader's relation to his past leader, I use Svolik [-@svolik_politics_2012]'s Authoritarian Spells, 1946 - 2008 data set that contains authoritarian leaders' affiliation with their previous leaders. An authoritarian leader is secure at time 0 when he is politically affiliated and from the same ruling coalition with his previous leader. An authoritarian leader is insecure at time 0 when a leader is in the opposition party of the previous party, openly opposes the previous leader, or he is unaffiliated (defined as not openly oppose or support) with the preceding government ([@svolik_politics_2012], p. 43). In other cases, if he is in a military regime or the regime with no authority, he is also defined as insecure when he starts his tenure. Note that in democracies, a leader's relation to his past is irrelevant; so there is no information of leader's relation to his past leader in democracies. All the democratic leaders are assumed as secure leaders before they serve in the office (at time *0*).

$\qquad$ The security of the regime over time *t* measures the vulnerability of the regime to collapse in any given year. I use political effectiveness score in the state fragility index to measure the political secureness of a regime to capture the dimensions of political opposition, citizen's confidence in the political process, political violence of a regime and related political risks [@marshall_global_2017]. The index ranges from 0 to 3, 0 means the most secure, and 3 means insecure. I recoded the regime's security at time *0* as a binary variable: those with score 0 in political securenss are coded secure, otherwise insecure. A worsening political effectiveness score over a leader's tenure captures a political upheaval that challenges the government authority. The wrosening score measures a shock to leader's security that causes instability. 

$\qquad$ Here I develop two types of shocks to security. Type (1): a leader who is presumably secure at time *0* becomes insecure immediately onward at time *1*. Substantively, it means a democratic leader or an authoritarian leader who politically affiliated with his previous leaders starts his tenure in an unstable and contested environment. The leaders may be highly constrained by the opposition and deep internal political divisions. Type (2): a leader experiences a shock to security during his tenure at time *t*. The political crisis captured in such a shock to security can be a result of threats from the fractionalized ruling coalition, popular uprisings, or even the actual use of force.  


$\qquad$ Here, I have developed the *independent variable* as a hypothesized treatment of shock to leader's security in the regime. It is a binary variable. Once a leader experienced either or both types of shocks to security, he is considered in the treatment group.  


$\qquad$ The *dependent variable* is a binary one. If a leader in the developing country has ever negotiated a South-North PTA with the provisions of economic reforms during his tenure, then this event happened, coded as 1, otherwise 0. Only the South-North preferential trade agreements that contain a competition chapter or a competition article is included. The information is collected on in DESTA data set [@durDesignInternationalTrade2014]. The provision includes but is not limited to the privatization of the state-owned firms and regulation on monopolies and cartels. A leader may negotiate a couple of PTAs, but only the first one will be only counted. In this data set, a majority of the PTAs have been put into force, while some of them were still in the negotiation process. Here I focus on the year leaders starts negotiation rather than the year of signatory because there may be leadership change between the years that a leader negotiates a PTA while another leader ratifies it. I collected the information on the state's initiation of PTA negotiations from 1995 to 2015.


$\qquad$ The unit of observation in the data set is leader. The data set covers 286 leaders clustered in 62 developing countries in the period 1995 to 2015. Note that in this data set, at least one of the leaders in these developing countries negotiated one PTA with the provision of competition policy with a developed country at some point during this period. The data set excludes leaders in the liberal democracies (V-Dem's Electoral Democracy Index above 0.75) where the rule of law and constraints on the executives are respected most of the time. In a regime as such, leadership change is routine and institutionalized. Therefore, there are lower risks for leaders losing power and hence lower incentives for leaders to use binding trade agreements for political survival. Note that under mature democratic regimes, all but the most extreme form of opposition will be channeled into the formal institutions of government, in which dissents of opposition can hardly dampen the political stability or cause a shock to security. In stable democracies, not only shocks to security rarely happen but also leader turnovers are generally institutionalized and peaceful. Leaders' reactions to such threats may also be institutionalized. Therefore, I limit the scope of this study in only autocratic and semi-democratic developing countries. Furthermore, it removes leaders whose tenure is less than one year, in such case they have no time to pursue any substantial policy changes given the short amount of time in office. 


$\qquad$ I completely aware that the hypothetical treatment assignment is not random in an observational study, so I include potential *confounding variables* to reduce ommited variable biases. The covariates in this study include the qualities of democratic or autocratic authority measured by V-Dem's Electoral Democracy Index [@coppedge_v-dem_2018] because state leaders face different institutional constraints when they initiate the trade agreement negotiation. State leaders also face different risks of shock to security in different regime types. The second covariate is leader's tenure and the length of uninterrupted regime duration up to a leader starts his tenure in Archigo data [@gleditsch_archigos_2015]. A leader's tenure is a confounder because the longer a leader holds office, the probability of engaging in the PTA negotiation may be higher, and the risks of being exposed to the shock to security is higher. 


$\qquad$ Two additional control variables are included to take into account the alternative hypothese. GDP per capita measured by the World Bank [@world_bank_world_2018] is a binary variable to capture economic recession of the developing countries between 1995 to 2015. In a period when a country experiences poor economic performance, a leader in the developing country perceives the potential economic benefits of the PTA with the South can help the economy to recover. Also, this regime may be more likely to experience negative shock of security when the economy declines. Government is more likely to lose legitimacy in times of economic difficulties; hence, it is more vulnerable to political risks. Therefore, there is an alternative possibility that a leader may negotiate a PTA driven by the economic benefits of the preferential trade agreement. 


$\qquad$ Human rights conditions measured by Political Terror Scale [@gibney_politicterror_2017] indicate the human rights practices of the regimes. A worsening Political Terror Scale measure an increased level of violence by the state engaging in state-sanctioned killings, torture, disappearances and political imprisonment. As illustrated above, a dominant understanding of leader's response after the opposition poses shock to security is to repress them. A worsening human rights conditions is a measurement of leader's punishment tactic to the opposition after he is challenged. If this competitive hypothesis holds, leaders will engage in more human rights violations after they experience the shock to security, a hypothetical treatment in this study.




```{r set up the data, echo=FALSE, message=FALSE, warning=FALSE, include=F}

getwd()
set_here(path=".", verbose=T)
here()

list.files(path=".")


```



```{r allregimesminusliberaldemocracy, results='hide', echo = FALSE, message = FALSE, warning=FALSE}

mydata_10_all <- read.csv("mydata10all_matching_3.csv") #286 observations
#View(mydata_10_all)

table(mydata_10_all$Z) #93 events
length(unique(mydata_10_all$ccode)) #62 countries

table(mydata_10_all$GDPdecline) #79 leaders experiencing decline

n_leader <- nrow(mydata_10_all) #286 leaders

#factorize Z
mydata_10_all$Z <- as.factor(mydata_10_all$Z)

percent_leader0 <- round((table(mydata_10_all$Z)[1]/n_leader)*100,2) #67.48%
percent_leader1 <- round((table(mydata_10_all$Z)[2]/n_leader)*100,2) #32.52%

percent_leader0 #leaders without shocks
percent_leader1 #leaders with shocks
```


```{r rearrange data, eval=T, echo=F, message=F, warning=F, results='hide'}

#########
#nondemocratic regimes:
mydata_10_auto <- subset(mydata_10_all, !DICT3=="democracy") #77 leaders in no authority plus autocracies

n_leader_auto <- nrow(mydata_10_auto)

percent_leader0_auto <- round((table(mydata_10_auto$Z)[1]/n_leader_auto)*100,2) #64%
percent_leader1_auto <- round((table(mydata_10_auto$Z)[2]/n_leader_auto)*100,2) #36%

percent_leader0_auto #leaders without shocks
percent_leader1_auto #leaders with shocks

##########
#democratic regimes:
mydata_10_dem <- subset(mydata_10_all, DICT3=="democracy") #209 in electoral democrcies (limited form of democracy)

n_leader_dem <- nrow(mydata_10_dem)

percent_leader0_dem <- round((table(mydata_10_dem$Z)[1]/n_leader_dem)*100,2) #69%
percent_leader1_dem <- round((table(mydata_10_dem$Z)[2]/n_leader_dem)*100,2) #31%

percent_leader0_dem #leaders without shocks
percent_leader1_dem #leaders with shocks


```


```{r all regime types, eval=T, echo=F, message=F, warning=F, results='hide', include=F}

dis_Z_event <- table(mydata_10_all$Z, mydata_10_all$event)
dis_Z_event[1,1] #142
dis_Z_event[1,2] #51
dis_Z_event[2,1] #52
dis_Z_event[2,2] #41


percent_leader0_event <- round(((dis_Z_event[1,2])/(dis_Z_event[1,2]+dis_Z_event[1,1]))*100,2) #26%
percent_leader1_event <- round(((dis_Z_event[2,2])/(dis_Z_event[2,1]+dis_Z_event[2,2]))*100,2) #44%


mydata10_all_ggplot2 <- na.omit(mydata_10_all[,c("Z", "event", "DICT3")])
mydata10_all_ggplot2$Z <- as.factor(mydata10_all_ggplot2$Z)

mydata10_all_ggplot2 <- ggplot(data=mydata10_all_ggplot2, aes(x=Z, y=event, color = DICT3))


mydata10_all_ggplot2  +
  geom_point(position=position_jitter(height=0.2, width=0.2)) + theme_bw() + 
  labs(x="Shock to Security", y = "Probability of PTAs Negotiation (%)") + 
  ggtitle("Dsitributions of Leaders across Regime Types \n Divided by Shock to Security") + 
  theme(plot.title = element_text(face="bold", size=16, hjust = 0.5)) + 
  annotate("text", x = 1:2, y = 0.5, label = c("26%", "44%")) +
  scale_x_discrete(breaks=c("0","1"), labels=c("No Shock", "With Shock"))

```


```{r distribution z event according to regime, eval=T, echo=F, message=F, warning=F, results='hide'}

dis_Z_event_auto <- table(mydata_10_auto$Z, mydata_10_auto$event)
dis_Z_event_auto[1,1] #142
dis_Z_event_auto[1,2] #51
dis_Z_event_auto[2,1] #52
dis_Z_event_auto[2,2] #41

percent_leader0_event_auto <- round(((dis_Z_event_auto[1,2])/(dis_Z_event_auto[1,2]+dis_Z_event_auto[1,1]))*100,2) #37%
percent_leader1_event_auto <- round(((dis_Z_event_auto[2,2])/(dis_Z_event_auto[2,1]+dis_Z_event_auto[2,2]))*100,2) #64%

percent_leader0_event_auto 
percent_leader1_event_auto


dis_Z_event_dem <- table(mydata_10_dem$Z, mydata_10_dem$event)
dis_Z_event_dem[1,1] #111
dis_Z_event_dem[1,2] #33
dis_Z_event_dem[2,1] #42
dis_Z_event_dem[2,2] #23

percent_leader0_event_dem <- round(((dis_Z_event_dem[1,2])/(dis_Z_event_dem[1,2]+dis_Z_event_dem[1,1]))*100,2) #23%
percent_leader1_event_dem <- round(((dis_Z_event_dem[2,2])/(dis_Z_event_dem[2,1]+dis_Z_event_dem[2,2]))*100,2) #35%

percent_leader0_event_dem 
percent_leader1_event_dem


```


```{r prepare table for percentages, eval = TRUE, echo = FALSE, results='hide', message = FALSE, warning = FALSE, include=FALSE}

#install.packages("xtable",repos = "http://cran.us.r-project.org")
library(xtable)

percent_Z <- c(percent_leader1, percent_leader1_dem, percent_leader1_auto)

percent_Z_event <- c(percent_leader1_event, percent_leader1_event_dem, percent_leader1_event_auto)

percentages <- as.matrix(rbind(percent_Z, percent_Z_event))

row_names_percentages <- c("Percentages of leaders being treated (%)", "Percentages of those treated negotiated a PTA (%)")
col_names_percentages <- c("All Regime Types", "Democracies", "Nondemocracies")
rownames(percentages) <- row_names_percentages
colnames(percentages) <- col_names_percentages

xtable(percentages)

```

```{r xtable print pvalues, results = "asis", message = FALSE, eval=T, echo=F, warning = FALSE}

knitr::kable(percentages, caption = "\\label{tab:per} Percentages of Leaders being Treated and those Trated Having an Event", format = "latex", booktabs = TRUE) %>%
          kable_styling(latex_options = "scale_down")
```


$\qquad$ Table \ref{tab:per} presents there are 93 events out of 286 total observations. Across regime types, `r percent_leader1`% of the leaders have experienced shocks to security. Among leaders across regimes, `r percent_leader1_event`% of those who have experienced a shock to security have chosen to negotiate a PTA with a developed country at some point in their tenure. `r percent_leader0_event`% of those who have not experienced a shock to security have negotiated a South-North PTA. Now, let us look at the relative frequency distribution divided by regime types. In non-democracies, `r percent_leader1_event_auto`% of leaders have experienced an insecure shock, while `r percent_leader1_dem`% in democracies have experienced one at some point during the time of their tenures. Among leaders in non-democracies regimes, `r percent_leader0_event_auto`% of those who have experienced a shock to security negotiated a South-North PTA. In contrast, only `r percent_leader1_dem`% of those who have experienced a shock to security in democracies negotiated one. Leaders in different regimes have different patterns of negotiating PTAs. The modeling approach adopted in this paper may allow us to account for such contextual factors.


#Model Analysis and Discussion

##Random Intercept Models

```{r prepare table for summary_s, eval = TRUE, echo = FALSE, results='hide', message = FALSE, warning = FALSE, include=FALSE}

#install.packages("xtable",repos = "http://cran.us.r-project.org")
library(xtable)

EVENT <- c("218", "0.32", "0.47", "0", "1")
Z <- c("218", "0.33", "0.47", "0", "1")
T_L <- c("218", "6.66", "6.51", "2", "45")
H_v <- c("218", "2.87","0.98","1", "5")
GDP <- c("49", "8.41", "0.92", "6.23", "11.02")
DEM <- c("49", "0.39", "0.19", "0.04", "0.74")
DEML <- c("49", "15.31", "13.04", "0", "42")
AUTL <- c("49", "9.22", "14.43", "0", "44")
REC <- c("49", "0.8", "0.41", "0", "1")

summary_s <- as.data.frame(rbind(EVENT, Z, T_L, H_v, GDP, DEM, DEML, AUTL, REC))


row_names_summary_s <- c("PTA Negotiation", "Shock to security", "Tenure of Leader", "Human rights violation", "GDP per capita (log)", "Level of democracy", "Age of democracy", "Age of autocracy", "Economic recession")
col_names_summary_s  <- c("N", "Mean", "SD", "Min.", "Max.")
rownames(summary_s) <- row_names_summary_s 
colnames(summary_s) <- col_names_summary_s 

xtable(summary_s)


```

```{r xtable print summary_s, results = "asis", message = FALSE, eval=T, echo=F, warning = FALSE}


library(knitr)
library(kableExtra)


knitr::kable(summary_s , caption = "\\label{tab:summary_s} Descriptive Statistics", format = "latex", booktabs = TRUE) %>%
          kable_styling(latex_options = "striped", full_width = F) %>%  kable_styling (latex_options = c("hold_position")) %>%    
    group_rows("Individual-Level", 1, 4) %>% group_rows("Group-Level", 5, 9) %>% landscape()

```

```{r prepare table for random intercept models, eval = TRUE, echo = FALSE, results='hide', message = FALSE, warning = FALSE, include=FALSE}

#install.packages("xtable",repos = "http://cran.us.r-project.org")
library(xtable)

model0_est <- c("-", "-", "-", "-", "-", "-", "-", "-", 
                "0.012", "0.208", "94.71%", "5.3%")
model1_est <- c("-", "-", "-", "-0.43", "0.00", "0.00", "-", "-", 
                "0.011", "0.2", "3.62%", "3.27%")
model1_p_value <- c("-", "-", "-", "0.09", "0.60", "0.14", "-", "-",  
                "-", "-", "-", "-")
model2_est <- c("-", "0.02", "-", "-0.64", "-", "-", "0.06", "-0.09",
                "0.016", "0.191", "7.03%", "-")
model2_p_value <- c("-", "0.017", "-", "0.03", "-", "-", "0.188", "0.395",
                "-", "-", "-", "-")
model3_est <- c("0.12", "0.19", "-0.08", "-0.58", "-", "-", "0.07", "-0.09",
                "0.014", "0.191", "8.05%", "-")
model3_p_value <- c("0.047", "0.017", "0.06", "0.004", "-", "-", "0.141", "0.39", "-", "-", "-", "-")

results <- as.data.frame(cbind(model0_est, model1_est, model1_p_value, model2_est, model2_p_value, model3_est, model3_p_value))


row_names_results <- c("Shock to security", "Tenure of Leader", "Human rights violation", "Level of democracy", "Age of democracy", "Age of autocracy", "GDP per capita (log)", "Economic recession", 
                       "Level-2 variance (between state)", "Level-1 variance (within states)", "Variation explained at level 1", "Variation explained at level 2")
col_names_results <- c(" ", "Estimates", "Pr(>|t|)", "Estimates", "Pr(>|t|)", "Estimates", "Pr(>|t|)")
rownames(results) <- row_names_results
colnames(results) <- col_names_results 

xtable(results)


```

```{r xtable print, results = "asis", message = FALSE, eval=T, echo=F, warning = FALSE}


library(knitr)
library(kableExtra)


knitr::kable(results, caption = "\\label{tab:re} Four Random Intercept Models of Explaining PTA Negotiation: Estimates, P-values and Variance Components", format = "latex", booktabs = TRUE) %>%
          kable_styling(latex_options = "striped", full_width = F) %>%  kable_styling (latex_options = c("hold_position")) %>%   add_header_above(c(" " = 1, "Model 0" = 1, "Model 1.1" = 2, "Model 1.2" = 2, "Model 1.3" = 2)) %>%  
    group_rows("Individual-Level", 1, 3) %>% group_rows("Group-Level", 4, 8) %>% group_rows ("Random Effects", 9, 12) %>% landscape()

```

$\qquad$ The data set in this paper exhibit a nested structure. The level-one units are state leaders, with a sample size of 286. The level-two units are 62 developing countries. Because of missing data, the sample size of level-one drops to 218, the sample size of level-two drops to 49. Table \ref{tab:summary_s} summarizes the descriptive statistics. If the models I use ignore the fact that individual leaders are clustered within different countries, I may run the risks of getting downwardly biased standard errors, which to inflated Type I errors. We may often use clustered standard errors to correct for the clustering data structure. Although it may be a solution of solving statistical problems, we may miss the opporunities to explore the theoretical questions of multilevel data. This research note will adopt a hierarchical linear modeling tehnique to answer the following question: After taking into account the differences of the regime types of the states, do individual leaders' shocks to security lead to leaders' decision of PTA negotiations? Results are shown in Table \ref{tab:re}.   


$\qquad$ The first step in building multilevel models is to decompose the variance in the dependent variable into two levels, the individual leader and state levels. This is equivalent to running a random intercept model with no explanatory variables, where I do not specify any predictors of PTA negotiation. It is often called an empty model. In decomposing the variance of dependent variable in this way, I can find the proportion of variance in the dependent variable attributable to between-cluster differences and the proportion attributable to within-cluster variability. I find that 5.3% of the variance in PTA negotiation explained at the country level, with the remaining 94.7% came from between-individual variation. Some variations at the state-level allows me to explore these contxtual variations in the multilevel model setting.     


$\qquad$ The second step is to build the multilevel models, starting from simple to a full and comlicated one. Only the level-two variables are added in the random intercept model. Three regime quality related state-level variables are added to see the effects of these variables on PTA negotiation. These three variables are the *age of democracy*, the *age of autocracy*, and the *level of democracy*. The second step is to include state-level covariates to see if countries with longer history of democracy or autocracy have positive association with the negotiate rate of preferential trade agreements. The reason lies in countries tend to form long-term relationship with trade partners and avoid transitioning and unstable countries. Also, when the developing countries' level of democracy increases, I expect there are more trade cooperation between these countries and developed countries that are most stable democracies. Reimge similarities may reward more trade opportunities among partners. Table \ref{tab:per} Model 1.1 shows the results. The ages of regimes show no substantive effects of PTA negotiation. The level of democracy shows the oppositie direction as I expected. All of these three variables have no statistical signifiance. This model has explained 3.27% of the level-2 variation captrued in the empty model.


$\qquad$ Model 1.2 keeps one of the level-2 variables used in Model 1.1, *level of democracy*, and adds in the variables that measure the economic conditions of states: *GDP per capita (logged)* and *economic recession*. A level-1 variable, the *tenure of leader* is also added in the model. This variable is centered within the groups. Note that the estimates of the two models for the three level-2 variables are the same, so I only show the result of the slightly complicated model with three level-2 variables and one level-1 variable. We can see level of democracy has statistical signifiance; yet, the estimate presents a negative relation between developing countries' level of democracy and PTA negotiation rates. The range of the level of democracy is from 0 to 1; if a regime's level of democracy increases from 0 to 1, suggesting an extreme transition from the most autocratic to the most democratic regime, the PTA negotiation rate goes down by 64%. More realistically, if a regime's level of democracy improves by 0.1 point in V-Dem, the probability of the PTA negotiation goes down by 6.4%. 


$\qquad$ I then add more level-1 variables, the *human rights conditions* and the main independent variable in this paper, the *shock to security*. The estimates in the random intercept models with and without the *shock to security* variable are similar, so I show the full model in Model 1.3. The variables *tenure of leader*, *GDP per capita(logged)* and *shock to security* have statisctical meanings, and the *human rights violation* is slightly above the threshold of rejecting the null hypothesis of no statistical significance (p-value 0.05). This full model only explains 8% of the level-1 variation identified in the empty model, which still leaves a large proportion of variation unexplained. Substantively, the full model shows holding other leader-related factors, political and economic conditions constant, when there is a political shock to security, the leader will increase the odds of PTA negotiation by 12%. When a leader holds office one year longer than the mean tenure years of leaders of that specific state, the leader will increase the probability of PTA negotiation by 2% when other factors are held constant.   








```{r Table results, results='asis', eval=F, echo = FALSE, message=FALSE, include=T}

library(stargazer)
#Now use stargazer to format the linear model
stargazer(lm3_all, header = FALSE, df = FALSE, omit = "fm2_p_all", intercept.bottom = F,
          title = "Ordinary Linear Squared Regression Model (After Full Matching): Results",
          covariate.labels = c("Constant", "Shock to Security", "Economic Recession","Mean Tenure"),
          dep.var.labels = "Preferential Trade Agreement Negotation Rate", 
          column.sep.width = "0.3pt", no.space = TRUE, font.size = "footnotesize", notes = "Automatic reporting p-values, need to be replaced with Fishers' Null")

```




$\qquad$ So far, I have created a hypothetical treatment, a shock to the leader's security, so leaders experienced such a shock are in treatment group, and those who did not are in control group. Because it is observational data, I use a statistical technique to balance the two groups so leaders being matched on the covariates are not too different in the treatment and control group. Then, I use an OLS model to estimate the probability that a leader in developing country engages in PTA at a given time during his tenure. The model is:
$$ Y = \beta_0 + \beta_1*ShocktoSecurity + \beta_2*EconomicRecession + \beta_3*Leaders'Tenures + \epsilon$$



#Preliminary Conclusion

Why and when will leaders in developing country negotiate South-North Preferential Trade Agreements? The answer is simple and intuitive: after leaders in developing country experience a negative shock that creates political instability, they are more likely to negotiate a PTA with the provision of economic reforms with the expectation to cut off the power sources of the disloyal opposition. Economic reform, hence, is a tactic for leader to punish the opposition and a toolkit to consolidate power in the regime. I have used original data on measuring security of leaders in developing countries between 1995 and 2015 to examine whether developing country leaders are more likely to negotiate a PTA in their tenures when they experience a political shock. I argue leaders in developing country approach the major liberal trade powers to negotiate a PTA with the provision of economic reform, so they have chosen the deepest PTAs by design. This is a strategic choice that leader makes with an objective to punish the opposition who defect from the promised support and pose the threatening demands. 




```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#References
